---
title: "Dynamics of Food Webs"
author: "Jonathan J. Borrelli"
date: "Tuesday, November 10, 2015"
output: ioslides_presentation
---
```{r, message = F, echo = F, warning = F}
library(knitr)
opts_chunk$set(tidy = T, message = F, warning = F, echo = F)
source("./Rscripts/MODELfunctions.R")
```

## Purpose
To further explore the relationship between food web structure and dynamics. 

## Hypothesis
Food web structure emerges from the preferential loss of unstable configurations. 

By simulating the dynamics of model food webs the structure of the food web following extinction events will exhibit a higher proportion of stable components via loss of species that contribute to destabilizing building blocks.  

## The Model

I have borrowed a model of multispecies predator prey dynamics based on species biomass modified from Yodzis and Innes (1992), that has been used in a number of publications. Specifically, I am using the form and parameterization suggested by Romanuk et al. (2009) and Williams and Martinez (2004). 

$$
\frac{dB_i(t)}{dt} = G_i(B) - x_iB_i(t) + \\ \sum_{j}^{n}\left(x_i y_{ij} F_{ij}(B) B_i(t) - \\ x_j y_{ji} F_{ji}(B) B_j(t) / e_{ji} \right)
$$

# Terms in the model 

## Growth

The first term, $G_i(B)$, is the function describing primary production of producer species in the absence of predation. Producers grow exponentially with density dependence. 

$$
G_i(B) = r_i B_i(t) (1 - \frac{B_i(t)}{K_i})
$$


In the equation, $r_i$ is the intrinsic rate of increase, $B_i$ is the biomass of population _i_, and $K_i$ is the carrying capacity of population _i_.

## Consumption

The consumption of resource _j_ by consumer _i_ is modeled by: 
$$
F_{ij}(B) = \frac{B_j^{1+q}}{\sum_k B_k^{1+q} + B_0^{1+q}}
$$

Here, $B_j$ is the biomass of the consumed resource, in the denominator the summed biomass across all $k$ resources, and $B_0$ is the half saturation density. The parameter $q$ is a tuning parameter that lets the modeller shift the functional response between a type II ($q = 0$) and a type III ($q = 1$). The functional response above defines the fraction of a predator's maximal ingestion that is realized at a given time step. 

## Functional Response

```{r FRsim, echo = F, cache = TRUE}
K = 1
x.i = .5
yij = 6
xpar = 0
B.o = .5
A =  matrix(c(0,0,1,0), nrow = 2)
FR = Fij

prey <- seq(0, 2, .01)
pred <- .5
x <- seq(0, 5, .2)
eaten <- matrix(nrow = length(prey), ncol = length(x))
for(j in 1:length(x)){
  for(i in 1:length(prey)){
    states = c(prey[i], pred)
    eaten[i,j] <- rowSums((x.i * yij * FR(states, A, B.o, xpar = x[j]) * states))[2]
  }
}
```

```{r FRfig, fig.height = 5, fig.width = 6, fig.cap = "How the functional response of the prey changes with altered values of the parameter q", echo = F}
matplot(prey, eaten, typ = "l", xlab = "Prey Biomass", ylab = "Prey Biomass Consumed")
```

## Parameter Values  


```{r params, echo = F, comment = NA, results = "asis"}
df <- data.frame(Parameter = factor(c("K", "x.i", "yij", "eij", "xpar", "B.o")), Value = c(1, .5, 6, 1, .2, .5))
kable(df, format = "pandoc", caption = "Fixed parameter values for the dynamic model")
```

# Model Run

## Create a network 


```{r,  fig.height = 6, fig.width = 6, fig.align='center'}
lay <- matrix(c(layout.sphere(graph.adjacency(nm1))[,1], TrophInd(nm1)$TL), ncol = 2)
plot(graph.adjacency(nm1), layout = lay)
```


## Get its properties

```{r, comment = NA, results = "asis"}
wp.in <- web_props(nm1)
kable(wp.in, format = "pandoc", caption = "Whole web properties of the initial niche model food web")
```

## Run Dynamics 

```{r, cache = T}
dyn2 <- Crmod(Adj = nm1, t = 1:100, G = G.i, method = conres, FuncRes = Fij, K = 1, x.i = .5, yij = 6, eij = 1, xpar = .2, B.o =.5, plot = F)

```

<iframe src="fwdyn.html"></iframe>


## Results

```{r dynamic-EX, echo = F, fig.width = 8, fig.height = 5, fig.cap = "Biomass dynamics of a niche model food web"}
matplot(dyn2[,-1], typ = "l", lwd = 2, xlab = "Time", ylab = "Biomass")
```

## Motifs

```{r}
load(file = "fw-test.Rdata")
ggplot(z2, aes(x = motif, y = value, fill = factor(q))) + geom_boxplot() + geom_line(aes(x = as.numeric(motif), y = 0)) + facet_wrap(~model, scales = "free_y")
```

## Other Structural Properties

```{r}
ggplot(r1[which(r1$property == "C"|r1$property == "apl"|r1$property == "top")], aes(x = initial, y = final, col = model)) + geom_point() + geom_abline(slope = 1, intercept = 0) + facet_wrap(~property, scales = "free")
```

